---
title: "Exploring Data with Visuals"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits = 2)
```

# Exploring the Video Games Dataset with Visuals

We started exploring the Video Games Dataset from Kaggle already, now let's add some analysis with visuals to get a better idea of what is going on with the data.

```{r environment_setup, echo=FALSE, error=TRUE, warning=TRUE}
library(ggplot2)
suppressMessages(library(dplyr))
suppressMessages(library(reshape2))
games = read.csv("../Datasets/video-game-sales-with-ratings/Video_Games_Sales_as_at_22_Dec_2016.csv", header=TRUE, na.strings = "")
```

# Box and Whisker Plots

These are a good way of visualising the median and first to third quartiles of our data. From the given dataset, we can see that the scores given to games lend themselves to a boxplot. Not too surprisingly, we can see that users (ie. game players) tend to be a little more generous in their scoring than critics are, which is probably why they are called critics.

```{r boxplot_scores, echo=TRUE}
scores = games %>% select(User_Score, Critic_Score)
scores$User_Score = as.numeric(scores$User_Score)
scores$Critic_Score = as.numeric(scores$Critic_Score)
scores = melt(scores, measure.vars = c("User_Score", "Critic_Score"), na.rm = TRUE) #melt to get variables right for boxplot
g = ggplot(scores, aes(variable, value))
g = g + stat_boxplot(geom = "errorbar") #add whiskers
g = g + geom_boxplot()  #add boxplot
g
```

## Histogram

Let's use histograms to show the disibution of video game sales per region.

```{r sales_histogram, echo=TRUE}

```